Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space

### 2025-04-17 paper/main.tex
Universal commits presents...

\title{Neural Networks Loss Landscape Convergence in Different Low-Dimensional Spaces}

@@

We consider a $K$-class classification problem. Let $\bx \in \bX$ be an input and $y \in \mathcal Y = \{1,\dots,K\}$ its label. A
neural network with parameters $\bw \in \bOmega \subset \mathbb R^P$ defines a mapping $f_{\bw} : \bX \to \mathbb R^K$. Given a dataset
$$
  \mathcal D =
  \{(\bx_i, y_i)\}_{i=1}^m
$$
of $m$ i.i.d. samples, and a per-sample loss $\ell \bigl( f_{\bw}(\bx_i), y_i \bigr)$ (e.g. cross–entropy), we define
$$
  \ell_i(\bw) =
  \ell \bigl( f_{\bw}(\bx_i) , y_i \bigr)\,.
$$

\begin{definition}
  The empirical loss on the first $k$ samples is
  $$
    \cL_k(\bw) =
    \frac1k \sum_{i=1}^k \ell_i(\bw),
  $$
  so that the full-sample loss is $\cL(\bw) = \cL_m(\bw)$.
\end{definition}

@@

\begin{definition}
  The Hessian of $\cL_k$ at $\bw$ is
  $$
    \bH_k(\bw) =
    \nabla^2_{\bw} \cL_k(\bw) =
    \frac1k \sum_{i=1}^k \nabla^2_{\bw} \ell_i(\bw),
  $$
  and, similarly, the full-sample Hessian is $\bH(\bw) = \bH_m(\bw) = \nabla^2_{\bw} \cL_m(\bw)$.
\end{definition}

@@

\begin{definition}
  Fix a tolerance $\Delta > 0$. The \emph{sufficient sample size} is
  $$
    k^* =
    \inf \{ k \mid \forall\, m \ge k : \; \Delta_m < \Delta\}.
  $$
  When $k \ge k^*$, adding further samples changes the loss landscape by less than $\Delta$.
\end{definition}

@@

\begin{assumption}
  The parameters $\bw$ are distributed according to a density $p(\bw)$ (e.g. a Gaussian prior centered at $\bw^*$).  Then
  $$
    \Delta_k =
    \int \bigl(\cL_{k+1}(\bw)-\cL_k(\bw)\bigr)^2\, p(\bw)\, d\bw =
    \mathbb{E}_{p(\bw)} \left( \cL_{k+1}(\bw) - \cL_{k}(\bw) \right)^2 =
  $$
  $$
    \mathbb D_{p(\bw)} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] +
    \Bigl( \mathbb E_{p(\bw)} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] \Bigr)^2.
  $$
\end{assumption}

@@

Concretely, let
$$
  \bP =
  \bigl[ \mathbf{e}_1, \dots, \mathbf{e}_d \bigr]
$$
be the matrix whose columns are the top $d$ eigenvectors.  We write any parameter vector as
$$
  \bw =
  \bw^* + \bP\, \btheta, \quad
  \btheta \in \mathbb R^d.
$$
Substituting into a second‑order Taylor expansion of $\cL_k$ around $\bw^*$ gives
$$
  \cL_k\bigl(\bw^* + \bP\,\btheta\bigr) \approx
  \cL_k(\bw^*) + \tfrac12\, \btheta^{\T}\! \bigl( \bP^{\T}\, \bH_k(\bw^*)\, \bP\bigr)\, \btheta =
  \cL_k(\bw^*) + \tfrac12\, \btheta^{\T} \bLambda_k\, \btheta,
$$
where
$$
  \bLambda_k =
  \bP^{\T}\, \bH_k(\bw^*)\, \bP =
  \operatorname{diag} \bigl( \lambda_k^1, \dots, \lambda_k^d \bigr).
$$
An identical expansion holds for $\cL_{k+1}$, yielding $\bLambda_{k+1}$ from the Hessian on $k+1$ samples.

\begin{theorem}[Approximate $\Delta_k$ via Eigenvalues]
  Under the assumption that $\btheta \sim \mathcal N(0, \sigma^2 \mathbf I_d)$, the change in loss can be estimated by
  $$
    \Delta_k =
    \mathbb E \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2 \approx
    \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i)^2
    + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^i - \lambda_k^i) \Bigr)^2 \Biggr).
  $$
\end{theorem}

\begin{remark}
  Assuming $\mathbb{E}_{p(\btheta)}(\btheta) = \mathbf{0}$ is natural when centering the sampling distribution at the minimum.
\end{remark}

@@

\begin{figure}[!htbp]
  \hspace*{-2.2cm}
  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_random_1_2.pdf}}
  \caption{Loss function landscape for small $k$, projected onto the two random directions.}
  \label{fig:loss_random_small}
\end{figure}

\begin{figure}[!htbp]
  \hspace*{-2.2cm}
  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_random_-2_-1.pdf}}
  \caption{Loss function landscape for maximum $k$, projected onto the two random directions.}
  \label{fig:loss_random_big}
\end{figure}

\paragraph{Loss Landscape for Small Sample Size}
In Figure~\ref{fig:loss_random_small} ($k=1$), the loss surface is highly irregular, with sharp peaks and deep valleys.
This irregularity indicates that with very few training samples, the model is extremely sensitive to small parameter perturbations.
The squared difference surface also exhibits large spikes, confirming that adding a single data point can cause substantial local changes
in the loss.

\paragraph{Loss Landscape for Maximum Dataset Size}
In contrast, Figure~\ref{fig:loss_random_big} (maximum $k$) shows a much smoother loss surface with gentle slopes. The overall loss
values are lower --- reflecting the averaging effect of more samples --- and the squared difference surface is correspondingly much smaller.
This smoothing demonstrates that as the dataset grows, the local optimum stabilizes, and additional samples produce only minor
adjustments to the loss landscape. Hence, beyond a certain dataset size, further data have negligible impact on the shape of the loss
surface.

@@

To study how the loss landscape converges as the dataset grows, we estimate and plot the function
$$
  \Delta_k =
  \mathbb{E}_{p(\bw)} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2
$$
via a Monte Carlo approximation using Gaussian perturbations in parameter space. The steps are:

\begin{enumerate}
  \item \textbf{Sample directions.}
        Starting from a fixed trained model with parameters $\bw^*$, draw $K$ independent random vectors
        $$
          \btheta^{(t)} \sim \mathcal{N} (\mathbf 0, \sigma^2 \mathbf I_d), \quad
          t = 1, \dots, K.
        $$

@@

  \item \textbf{Compute Monte Carlo estimate.}
        Approximate
        $$
          \Delta_k \approx
          \frac{1}{K} \sum_{t=1}^K d_t.
        $$
        Increasing $K$ reduces the variance of this estimator.

@@

\begin{figure}[!htbp]
  \hspace*{-2.6cm}
  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_sigma_2_64.pdf}}
  \caption{Delta function landscape varying the Gaussian variance $\sigma$.}
  \label{fig:delta_random_sigma}
\end{figure}

\begin{figure}[!htbp]
  \hspace*{-2.6cm}
  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_dim_1_64.pdf}}
  \caption{Delta function landscape varying the number of dimensions.}
  \label{fig:delta_random_dim}
\end{figure}

@@

\paragraph{Projection Directions.}
In the main experiment, we replace random directions with the $d=10$ eigenvectors of the Hessian at the local minimizer $\bw^*$ that
correspond to its largest eigenvalues.  Prior work \cite{sagun2018empirical} shows these top modes capture nearly all of the
non‑negligible curvature, and Figure~\ref{fig:evgen} confirms a rapid eigenvalue decay. Fixing $d=10$ balances computational cost and
fidelity, and our tests on Fashion‑MNIST and CIFAR demonstrate that this choice suffices even for complex architectures.

@@

Figures~\ref{fig:loss_eigen_small} and \ref{fig:loss_eigen_big} display 3D plots of $\cL_k(\btheta)$ (left) and $\bigl[\cL_{k+1}(\btheta)-\cL_k(\btheta)\bigr]^2$ (right), where $\btheta$ varies over the span of the top two Hessian eigenvectors.  The procedure mirrors the random‑direction case but uses these maximally informative axes.

\begin{figure}[!htbp]
  \hspace*{-2.2cm}
  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_eigen_1_2.pdf}}
  \caption{Loss landscape $\cL_k$ and squared difference for small $k$, projected onto the top two Hessian eigenvectors.}
  \label{fig:loss_eigen_small}
\end{figure}

\begin{figure}[!htbp]
  \hspace*{-2.2cm}
  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_eigen_-2_-1.pdf}}
  \caption{Loss landscape $\cL_k$ and squared difference for maximum $k$, projected onto the top two Hessian eigenvectors.}
  \label{fig:loss_eigen_big}
\end{figure}

In Figure~\ref{fig:loss_eigen_small} (small $k$, e.g.\ $k=1$), the loss surface projected onto the top two Hessian eigenvectors is
noticeably more structured and less erratic than in the random‐direction case. This confirms that these eigenvectors capture the
most influential directions of loss variation. The squared‐difference plot also shows larger peaks, indicating that the model is
highly sensitive to new data along these key axes when $k$ is small.

By contrast, Figure~\ref{fig:loss_eigen_big} (maximum $k$) shows a much smoother loss surface and a greatly reduced squared‐difference.
As the dataset grows, the loss changes less along the principal curvature directions, and the model’s local behavior stabilizes.
Notably, even at large $k$, the squared difference remains higher than in the random‐direction experiment, providing a practical upper
bound on landscape fluctuations. These findings reinforce that projecting onto the top eigenvectors effectively reveals the convergence
behavior of the loss landscape and helps intuitively understand the existence of the point beyond which additional data have minimal impact.

@@

\begin{figure}[!htbp]
  \hspace*{-2.6cm}
  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_eigen_1_10_64.pdf}}
  \caption{Delta function landscape}
  \label{fig:delta_eigen}
\end{figure}

@@

We approximate $\Delta_k$ by
$$
  \Delta_k \approx
  \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^{d} (\lambda_{k+1}^i - \lambda_{k}^i)^2 +
  \Bigl( \sum_{i=1}^{d} (\lambda_{k+1}^i - \lambda_{k}^i) \Bigr)^2 \Biggr),
$$
where $\{\lambda_k^i\}_{i=1}^d$ and $\{\lambda_{k+1}^i\}_{i=1}^d$ are the top $d$ Hessian eigenvalues at sample sizes $k$ and $k + 1$,
and $\sigma$ is the perturbation variance.

@@

\begin{figure}[!htbp]
  \subfloat{\includegraphics[width=1\textwidth]{img/delta_border_1_10_64.pdf}}
  \caption{Delta function landscape}
  \label{fig:delta_eigen_comparison}
\end{figure}

@@

Our Hessian‐projection framework provides a clear picture of how a neural network’s loss landscape evolves with training–set size.
By restricting attention to the $d$ principal curvature directions (the top Hessian eigenvectors), we obtain far more structured 3D
loss surfaces and a sharply decaying landscape‐difference measure
$$
  \Delta_k \;=\; \mathbb{E}_{p(\btheta)}\bigl[L_{k+1}(\btheta)-L_k(\btheta)\bigr]^2,
$$
which empirically follows an approximate $O(1/k)$ trend.  Moreover, the Hessian‐based bound
$$
  \Delta_k\approx\frac{\sigma^4}{4}\Bigl(2\sum_{i=1}^d(\lambda^i_{k+1}-\lambda^i_k)^2
  +\bigl(\sum_{i=1}^d(\lambda^i_{k+1}-\lambda^i_k)\bigr)^2\Bigr)
$$
consistently upper‐bounds the Monte Carlo estimates (\ref{fig:delta_eigen_comparison}). This confirms that only a small number of
eigen‐modes drive the bulk of loss‐landscape change.

>>>>

\title{Neural Networks Loss Landscape Convergence in Hessian Low-Dimensional Space}

@@

We consider a $S$-class classification problem. Let $\bx \in \bX$ be an input and $y \in \mathcal Y = \{1, \dots, S\}$ its label. A
neural network with parameters $\bw \in \bOmega \subset \mathbb R^P$ defines a mapping $f_{\bw} : \bX \to \mathbb R^S$. Given a dataset
$$
  \mathcal D =
  \{(\bx_i, y_i)\}_{i=1}^N
$$
of $m$ i.i.d. samples, and a per-sample loss $\ell \bigl( f_{\bw}(\bx_i), y_i \bigr)$ (e.g. cross–entropy), we define
$$
  \ell_i(\bw) =
  \ell \bigl( f_{\bw}(\bx_i) , y_i \bigr)\,.
$$

\begin{definition}
  The empirical loss on the first $k$ samples is
  $$
    \cL_k(\bw) =
    \frac1k \sum_{i=1}^k \ell_i(\bw),
  $$
  so that the full-sample loss is $\cL(\bw) = \cL_N(\bw)$.
\end{definition}

@@

\begin{definition}
  The Hessian of $\cL_k$ at $\bw$ is
  $$
    \bH_k(\bw) =
    \nabla^2_{\bw} \cL_k(\bw) =
    \frac1k \sum_{i=1}^k \nabla^2_{\bw} \ell_i(\bw),
  $$
  and, similarly, the full-sample Hessian is $\bH(\bw) = \bH_N(\bw) = \nabla^2_{\bw} \cL_N(\bw)$.
\end{definition}

@@

\begin{definition}
  Fix a tolerance $\Delta > 0$. The \emph{sufficient sample size} is
  $$
    K =
    \inf \{ k \mid \forall\, k' \ge k : \; \Delta_{k'} < \Delta\}.
  $$
  When $k \ge K$, adding further samples changes the loss landscape by less than $\Delta$.
\end{definition}

@@

\begin{assumption}
  The parameters $\bw$ are distributed according to a density $p(\bw)$ (e.g. a Gaussian prior centered at $\bw^*$).  Then
  $$
    \Delta_k =
    \int \bigl(\cL_{k+1}(\bw)-\cL_k(\bw)\bigr)^2\, p(\bw)\, d\bw =
    \mathbb{E} \left( \cL_{k+1}(\bw) - \cL_{k}(\bw) \right)^2 =
  $$
  $$
    \mathbb D \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] +
    \Bigl( \mathbb E \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr] \Bigr)^2.
  $$
\end{assumption}

@@

Concretely, let
$$
  \bP =
  \bigl[ \mathbf{e}_1, \dots, \mathbf{e}_d \bigr]
$$
be the matrix whose columns are the top $d$ eigenvectors.  We write any parameter vector as
$$
  \bw =
  \bw^* + \bP\, \btheta, \quad
  \btheta \in \mathbb R^d.
$$
Substituting into a second‑order Taylor expansion of $\cL_k$ around $\bw^*$ gives
$$
  \cL_k\bigl(\bw^* + \bP\,\btheta\bigr) \approx
  \cL_k(\bw^*) + \tfrac12\, \btheta^{\T}\! \bigl( \bP^{\T}\, \bH_k(\bw^*)\, \bP\bigr)\, \btheta =
  \cL_k(\bw^*) + \tfrac12\, \btheta^{\T} \bLambda_k\, \btheta,
$$
where
$$
  \bLambda_k =
  \bP^{\T}\, \bH_k(\bw^*)\, \bP =
  \operatorname{diag} \bigl( \lambda_k^{(1)}, \dots, \lambda_k^{(d)} \bigr).
$$
An identical expansion holds for $\cL_{k+1}$, yielding $\bLambda_{k+1}$ from the Hessian on $k+1$ samples.

\begin{theorem}[Approximate $\Delta_k$ via Eigenvalues]
  Under the assumption that $\btheta \sim \mathcal N(0, \sigma^2 \mathbf I_d)$, the change in loss can be estimated by
  $$
    \Delta_k =
    \mathbb E \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2 \approx
    \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)})^2
    + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)}) \Bigr)^2 \Biggr).
  $$
\end{theorem}

\begin{remark}
  Assuming $\mathbb{E}(\btheta) = \mathbf{0}$ is natural when centering the sampling distribution at the minimum.
\end{remark}

@@

\begin{figure}[!htbp]
  \hspace*{-2.2cm}
  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_random_1_2.pdf}}
  \caption{$\cL_k$ (left) and $\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right) for small $k$, projected onto the two random vectors.}
  \label{fig:loss_random_small}
\end{figure}

\begin{figure}[!htbp]
  \hspace*{-2.2cm}
  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_random_-2_-1.pdf}}
  \caption{$\cL_k$ (left) and $\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right) for maximum $k$, projected onto the two random vectors.}
  \label{fig:loss_random_big}
\end{figure}

\paragraph{Loss Landscape for the Small Dataset}
In Figure~\ref{fig:loss_random_small}, the loss surface is highly irregular, with sharp peaks and deep valleys.
This irregularity indicates that with very few training samples, the model is extremely sensitive to small parameter perturbations.
The squared difference surface also exhibits large spikes, confirming that adding a single data point can cause substantial local changes
in the loss.

\paragraph{Loss Landscape for the Full Dataset}
In contrast, Figure~\ref{fig:loss_random_big} shows a much smoother loss surface with gentle slopes. The overall loss
values are lower --- reflecting the averaging effect of more samples --- and the squared difference surface is correspondingly much smaller.
This smoothing demonstrates that as the dataset grows, the local optimum stabilizes, and additional samples produce only minor
adjustments to the loss landscape. Hence, beyond a certain dataset size, further data have negligible impact on the shape of the loss
surface.

@@

To study how the loss landscape converges as the dataset grows, we estimate and plot the function
$$
  \Delta_k =
  \mathbb{E} \bigl[ \cL_{k+1}(\bw) - \cL_k(\bw) \bigr]^2
$$
via a Monte Carlo approximation using Gaussian perturbations in parameter space. The steps are:

\begin{enumerate}
  \item \textbf{Sample directions.}
        Starting from a fixed trained model with parameters $\bw^*$, draw $T$ independent random vectors
        $$
          \btheta^{(t)} \sim \mathcal{N} (\mathbf 0, \sigma^2 \mathbf I_d), \quad
          t = 1, \dots, T.
        $$

@@

  \item \textbf{Compute Monte Carlo estimate.}
        Approximate
        $$
          \Delta_k \approx
          \frac{1}{T} \sum_{t=1}^T d_t.
        $$
        Increasing $K$ reduces the variance of this estimator.

@@

\begin{figure}[!htbp]
  \hspace*{-2.6cm}
  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_sigma_2_64.pdf}}
  \caption{$\Delta_k$ (left) and $\Delta_k \cdot k^2$ (right), varying the Gaussian variance with fixed $d = 2$.}
  \label{fig:delta_random_sigma}
\end{figure}

\begin{figure}[!htbp]
  \hspace*{-2.6cm}
  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_random_dim_1_64.pdf}}
  \caption{$\Delta_k$ (left) and $\Delta_k \cdot k^2$ (right), varying the subspace dimension with fixed $\sigma = 1$.}
  \label{fig:delta_random_dim}
\end{figure}

@@

\paragraph{Projection Directions.}
In the main experiment, we replace random directions with the $d$ eigenvectors of the Hessian at the local minimizer $\bw^*$ that
correspond to its largest eigenvalues.  Prior work \cite{sagun2018empirical} shows these top modes capture nearly all of the
non‑negligible curvature, and Figure~\ref{fig:evgen} confirms a rapid eigenvalue decay. Fixing $d=10$ balances computational cost and
fidelity, and our tests on Fashion‑MNIST and CIFAR demonstrate that this choice suffices even for complex architectures.

@@

Figures~\ref{fig:loss_eigen_small} and \ref{fig:loss_eigen_big} display 3D plots of $\cL_k(\btheta)$ (left) and 
$\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right), where $\btheta$ varies over the span of the top two Hessian eigenvectors.
 The procedure mirrors the random‑direction case but uses these maximally informative axes.

\begin{figure}[!htbp]
  \hspace*{-2.2cm}
  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_eigen_1_2.pdf}}
  \caption{$\cL_k$ (left) and $\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right) for small $k$, projected onto the top two Hessian eigenvectors.}
  \label{fig:loss_eigen_small}
\end{figure}

\begin{figure}[!htbp]
  \hspace*{-2.2cm}
  \subfloat{\includegraphics[width=1.3\textwidth]{img/loss_eigen_-2_-1.pdf}}
  \caption{$\cL_k$ (left) and $\bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2$ (right) for maximum $k$, projected onto the top two Hessian eigenvectors.}
  \label{fig:loss_eigen_big}
\end{figure}

\paragraph{Loss Landscape for the Small Dataset}
In Figure~\ref{fig:loss_eigen_small}, the loss surface projected onto the top two Hessian eigenvectors is
noticeably more structured and less erratic than in the random‐direction case. This confirms that these eigenvectors capture the
most influential directions of loss variation. The squared‐difference plot also shows larger peaks, indicating that the model is
highly sensitive to new data along these key axes when $k$ is small.

\paragraph{Loss Landscape for the Full Dataset}
By contrast, Figure~\ref{fig:loss_eigen_big} shows a much smoother loss surface and a greatly reduced squared‐difference.
As the dataset grows, the loss changes less along the principal curvature directions, and the model’s local behavior stabilizes.
Notably, even at large $k$, the squared difference remains higher than in the random‐direction experiment, providing a practical upper
bound on landscape fluctuations. These findings reinforce that projecting onto the top eigenvectors effectively reveals the convergence
behavior of the loss landscape and helps intuitively understand the existence of the point beyond which additional data have minimal impact.

@@

\begin{figure}[!htbp]
  \hspace*{-2.6cm}
  \subfloat{\includegraphics[width=1.35\textwidth]{img/delta_eigen_1_10_64.pdf}}
  \caption{$\Delta_k$ (left) and $\Delta_k \cdot k^2$ (right)}
  \label{fig:delta_eigen}
\end{figure}

@@

We approximate $\Delta_k$ by
$$
  \Delta_k \approx
  \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)})^2
  + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)}) \Bigr)^2 \Biggr).
$$
where $\{\lambda_k^{(i)}\}_{i=1}^d$ and $\{\lambda_{k+1}^{(i)}\}_{i=1}^d$ are the top $d$ Hessian eigenvalues at sample sizes $k$ and $k + 1$,
and $\sigma$ is the perturbation variance.

@@

\begin{figure}[!htbp]
  \subfloat{\includegraphics[width=1\textwidth]{img/delta_border_1_10_64.pdf}}
  \caption{Theoretical and Empirical $\Delta_k$}
  \label{fig:delta_eigen_comparison}
\end{figure}

@@

Our Hessian‐projection framework provides a clear picture of how a neural network’s loss landscape evolves with training–set size.
By restricting attention to the $d$ principal curvature directions (the top Hessian eigenvectors), we obtain far more structured 3D
loss surfaces and a sharply decaying landscape‐difference measure
$$
  \Delta_k =
  \mathbb{E} \bigl[ \cL_{k+1}(\btheta) - \cL_k(\btheta) \bigr]^2,
$$
which empirically follows an approximate $O(1/k)$ trend.  Moreover, the Hessian‐based bound
$$
  \Delta_k \approx
  \frac{\sigma^4}{4}\, \Biggl( 2 \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)})^2
  + \Bigl( \sum_{i=1}^d (\lambda_{k+1}^{(i)} - \lambda_k^{(i)}) \Bigr)^2 \Biggr).
$$
consistently upper‐bounds the Monte Carlo estimates (\ref{fig:delta_eigen_comparison}). This confirms that only a small number of
eigen‐modes drive the bulk of loss‐landscape change.
